---
title: "DEV REPORT `r params$title`"
subtitle: "`r params$subtitle`"
author: "`r params$author`"
format: 
  html:
    fig-width: 12
    fig-height: 10
    toc: true
    toc-title: 'N-term Report'
    title-block-banner: '#1b2944'
    title-block-banner-color: '#f79646'
    lightbox: auto

execute: 
  echo: false
  warning: false
embed-resources: true
theme: 
 - layout/brand 
 - layout/custom.scss

params:
  description: ""
  design_file: ""
  input_file: ""
  folder_prj: ""
  title: ""
  subtitle: ""
  author: "" 
  select_group: ""
  sample_stat: ""
  grp_stat: ""
  glb_stat: ""
  pep_id: ""
  pep_id_group: ""
  ace_group: ""
  ace_: ""
 

---


```{css css_set_up}
/* Custom CSS for better visualization */
.card {
  border: 1px solid #dee2e6;
  border-radius: 0.25rem;
  margin-bottom: 1rem;
}

.card-header {
  background-color: #f8f9fa;
  border-bottom: 1px solid #dee2e6;
  padding: 0.75rem 1.25rem;
}

.card-body {
  padding: 1.25rem;
}

.badge {
  font-size: 0.9em;
  padding: 0.5em 1em;
}

.elements-table {
  margin-top: 1rem;
}

.dataTables_wrapper {
  padding: 10px;
}
.card-header {
  transition: background-color 0.3s ease;
}

.card-header:hover {
  background-color: #e9ecef;
}

.collapse-icon {
  transition: transform 0.3s ease;
}

.card-header[aria-expanded="true"] .collapse-icon {
  transform: rotate(180deg);
}

/* Animate collapse */
.collapse {
  transition: height 0.35s ease;
}

/* Optional: Style to show which card is active */
.card-header[aria-expanded="true"] {
  background-color: #e9ecef;
  border-bottom: none;
}

/* Add spacing between header elements */
.card-title {
  margin-bottom: 0;
}

.card-subtitle {
  margin-top: 0.5rem;
}

/* Ensure the chevron icon aligns properly */
.fas {
  width: 1em;
  text-align: center;
}
```

```{js js_set_up}
// Add this JavaScript to handle the chevron rotation
document.addEventListener('DOMContentLoaded', function() {
  document.querySelectorAll('.card-header').forEach(header => {
    header.addEventListener('click', function() {
      const isExpanded = this.getAttribute('aria-expanded') === 'true';
      this.setAttribute('aria-expanded', !isExpanded);
    });
  });
  
  // Show first card by default (optional)
  const firstCard = document.querySelector('.collapse');
  if (firstCard) {
    new bootstrap.Collapse(firstCard, {
      toggle: false
    }).show();
  }
});
```

```{r import_data}

L <- readLines(params$design_file, n = 1)
    if (grepl(";", L)) expdesign <- read.csv2(params$design_file) else expdesign <- read.csv(params$design_file)

glb_df  <-  readRDS(params$glb_stat )
grp_df <-  readRDS(params$grp_stat )

pepid_df  <-  readRDS(params$pep_id )
pepidgrp_df <-  readRDS(params$pep_id_group )

l_ace_grp <-  readRDS(params$ace_group )
l_ace_ <-  readRDS(params$ace_ )



```

```{r load_packg_set_up}
#| output: false
 required_packages = c("ggrepel","ggplot","reactable",
                       "plotly","DT",'gridExtra','ggVennDiagram','dplyr' ,'logger','pander','yaml','upsetjs','htmltools','purrr')
 ntermreport::check_dependencies(required_packages)

 st=format(Sys.time(), "%Y%m%d_%H%M")
write(as.yaml(params), file =  file.path( params$folder_prj, paste0("NtermDDA_parameter_",st  , ".yaml")))

theme_custom_vis <- function(base_size = 12) {
  theme_bw(base_size = base_size) %+replace%
    theme(
      # leggend
      legend.title = element_text(size = rel(0.85), face = "bold"),
      legend.text = element_text(size = rel(0.70), face = "bold"),
      legend.key = element_rect(fill = "transparent", colour = NA),
      legend.key.size = unit(1.5, "lines"),
      legend.background = element_rect(fill = "transparent", colour = NA),
      strip.background = element_rect(fill = "#bababa", color = "black"),
      strip.text = element_text(size = rel(0.85), face = "bold", color = "#1b2944", margin = margin(5,0,5,0))
    )
}

theme_set(theme_custom_vis(base_size = 18))

if (all(params$select_group != '')) {
  log_info(paste0('Selected groups: ',paste(params$select_group,collapse = ' '  )))
  selec_group_l <-  params$select_group

  all_group_ <- unique(expdesign$Group)
  }


## reactable theme 

lab_report_theme <- reactableTheme(
  color = "#2c3e50",
  backgroundColor = "#ffffff",
  borderColor = "#dfe2e5",
  borderWidth = "1px",
  
  # Styling the Header
  headerStyle = list(
    backgroundColor = "#f8f9fa",
    color = "#34495e",
    fontWeight = "700",
    borderBottom = "2px solid #3498db" # Adds a nice blue accent line
  ),
  
  # Styling the Table Body
  rowStyle = list(
    fontSize = "18px"
  ),
  
  # Styling the Search Box (if you use search = TRUE)
  searchInputStyle = list(
    width = "100%",
    backgroundColor = "#fdfdfd"
  )
)


```

# Experiment Description

`r params$description`

#  General PSMs Count
 
```{r general_viz}

#  reactable(glb_df,  columns = list(
#    percentage = colDef(format = colFormat(percent = FALSE, digits = 2,suffix='%'))
#    )
#  )

metric_descriptions <- c(
  "N-terminally"           = "PSMs identified with a free or modified N-terminus.",
  "Ace"                     =     "PSMs N-terminally Acetilated",    
  "C-terminal"             = "PSMs representing the protein C-terminus.",
  'Gln-NtermQ'              = "PSMs N-termilayy with  pyro-glutamate modification",
  "NH2"                    = "PSMs with C-terminal.",
  "N-terminally with H"     = 'PSMs N-terminally that contain Histidine in their sequence',
  "%.enrich.SCX.step"      = "Efficiency of the Strong Cation Exchange enrichment step.",
  "Total PSM"              = "Sum of N-terminally, C-terminal, and NH2 PSMs.",
  "Nterminally Acetylated" = "Fraction of N-terminal PSMs carrying an Acetyl group."
)
reactable(
  glb_df, 
  pagination = FALSE,
  highlight = TRUE,
  rowStyle = function(index) {
    if (glb_df$label[index] == "Total PSM") {
      list(fontWeight = "bold", background = "#f8f9fa", borderTop = "2px solid #dee2e6")
    } else if (glb_df$label[index] == "%.enrich.SCX.step") {
      list(fontStyle = "italic", color = "#2c3e50")
    }
  },
  columns = list(
    # Use 'Metric' if you renamed the column as discussed
    label = colDef(
      name = "Label",
      cell = function(value) {
        # Lookup description based on the row value
        desc <- metric_descriptions[value]

        # 2. Force it to be a simple string (remove names and handle NULLs)
        if (is.na(desc) || is.null(desc)) {
          display_text <- ""
        } else {
          display_text <- unname(desc) 
        }

        # Return an HTML span with a title attribute for the tooltip
        tags$span(style = "cursor: help; border-bottom: 1px dotted #ccc;", 
                  title = display_text, 
                  value)
      }
    ),
    count_absolute = colDef(name = "Count", format = colFormat(separators = TRUE,  locales = "fr-FR")),
    percentage = colDef(
      name = "Percentage",
      format = colFormat(digits = 2, suffix = "%")
    )
  ),
  theme = lab_report_theme
)
```



## PSMs Count per Sample

```{r grp_viz}
grp_df <- grp_df %>%  arrange(sample)


nACe_pec <-  grp_df %>% filter(label == 'Nterminally Acetylated') %>% arrange(sample)

summary_stats <- nACe_pec %>%
  summarise(
    Metric = "N-terminally Acetylated (%)",
    n_samples = n(),
    Mean = mean(percentage, na.rm = TRUE),
    SD = sd(percentage, na.rm = TRUE),
    CV_percent = (SD / Mean) * 100  # Coefficient of Variation is often useful in PTM analysis
  )
ace_mean <- mean(nACe_pec$percentage, na.rm = TRUE)
ace_sd <- sd(nACe_pec$percentage, na.rm = TRUE)

Nter_ace_perc_plot <-   ggplot ( nACe_pec, aes(x = factor(sample, levels = sample), y = percentage)) +
  geom_col(fill = "steelblue") +
  # Add the mean line
  geom_hline(yintercept = ace_mean, linetype = "dashed", color = "red", size = 1) +
  # Add a label for the mean
  annotate("text", x = -Inf, y = ace_mean + 1, label = paste("Mean:", round(ace_mean, 2)), color = "red", hjust = - 0.1) +
  scale_x_discrete(drop = FALSE) +
  labs(x = "Sample", y = "% Nterminally Acetylated", title = "Nterminally Acetylated") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

Nter_ace_perc_plot

##
reactable(
  summary_stats,
  columns = list(
    Metric = colDef(name = "Analysis Target", minWidth = 200, style = list(fontWeight = "bold")),
    n_samples = colDef(name = "Sample Count", align = "center"),
    Mean = colDef(
      name = "Average (%)",
      format = colFormat(digits = 2, suffix = "%"),
      style = list(background = "#f0f9ff", fontWeight = "bold")
    ),
    SD = colDef(
      name = "Std. Deviation",
      format = colFormat(digits = 2)
    ),
    CV_percent = colDef(
      # We replace 'name' and 'helpText' with a custom header function
      header = function(value) {
        tags$span(title = "Coefficient of Variation (SD/Mean)", "CV (%)")
      },
      format = colFormat(digits = 1, suffix = "%")
    )
  ),
  bordered = TRUE,
  highlight = TRUE,
  theme = lab_report_theme

)



## here should do the mean visualization 

 reactable(grp_df, groupBy= 'sample' ,  
 pagination = FALSE,
  highlight = TRUE,
  rowStyle = function(index) {
    row_label <- grp_df$label[index]
    if (is.na(row_label)) return(NULL) # Skip styling for the group header rows
    if (row_label == "Total PSM") {
      list(fontWeight = "bold", background = "#f8f9fa", borderTop = "2px solid #dee2e6")
    } else if (row_label == "%.enrich.SCX.step") {
      list(fontStyle = "italic", color = "#2c3e50")
    }
    },
  columns = list(
    # Use 'Metric' if you renamed the column as discussed
    label = colDef(
      name = "Label",
      cell = function(value) {
        # Lookup description based on the row value
        desc <- metric_descriptions[value]

        # 2. Force it to be a simple string (remove names and handle NULLs)
        if (is.na(desc) || is.null(desc)) {
          display_text <- ""
        } else {
          display_text <- unname(desc) 
        }

        # Return an HTML span with a title attribute for the tooltip
        tags$span(style = "cursor: help; border-bottom: 1px dotted #ccc;", 
                  title = display_text, 
                  value)
      }
    ),
    count_absolute = colDef(name = "Count", format = colFormat(separators = TRUE, locales = "fr-FR")),
    percentage = colDef(
      name = "Percentage",
      format = colFormat(digits = 2, suffix = "%")
    )
  ),
  theme = lab_report_theme

)
 
```


# Quantitative QC 


Here goes other text.  (From Peptide file )

```{r idplot}
#| layout-ncol: 2
#| fig-cap: 
#|   - "Quantified precursor for all the samples"
#|   - "Quantified precursor aggregated per groups"

fig.prec <-
  ggplot (pepid_df, aes(x = factor(Sample, levels = Sample), y = n_precursors)) +
  geom_col (fill = "steelblue") +
  scale_x_discrete(drop = FALSE) +
  labs(x = "Sample", y = "Number of modified peptides", title = "Unique modified peptides") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

fig.prec




fig.grp <-
  ggplot (pepidgrp_df, aes(x = factor(Group, levels = Group), y = n_precursors)) +
  geom_col (fill = "steelblue") +
  scale_x_discrete(drop = FALSE) +
  labs(x = "Group", y = "Number of modified peptides", title = "Unique modified peptides") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

fig.grp

```


#  Overlapping Acetylated Peptide Across Groups


*To be checked !!*
I use all the precursor (from PEPTIDE file) with: 

- *Nterm_mod.HS == "Acetyl* and *quant_valid.HS == TRUE*

- *pep_start is 1|2*.

for every group I can make an upset plot, containing the replicates. See below.

```{r upset_peptide}
#| results: asis

#pepidsample_df

grp_value  <- expdesign %>% distinct(Group) %>% pull()
my_list <- list()
my_peptide_info <- list()
for ( g_ in grp_value) {
      my_list[[g_]] <- l_ace_grp[[g_]]$pep_seq
            my_peptide_info[[g_]] <-  l_ace_grp[[g_]] %>% select(prot_acc,pep_seq,prot_desc)

}

ann_df <- bind_rows(my_peptide_info) %>% distinct(prot_acc, pep_seq,prot_desc)

ntermreport::render_child(data = my_list ,
              label = g_ , 
              path= "",  
              template='_template_upsetDIAG.Rmd') 

#upsetjs() %>%  fromList(my_list )  %>% generateDistinctIntersections()

all_elements <- unique(unlist(my_list)) 
all_elements <- all_elements[!is.na(all_elements)]
# Build logical membership table
intersection_ <- data.frame(
  Elements = all_elements,
  sapply(names(my_list), function(nm) all_elements %in% my_list[[nm]])
)

all_intersections_ <- ntermreport::get_all_intersections(intersection_)

all_intersections_filt <- all_intersections_[sapply(all_intersections_, function(x) x$count > 50)]

cat('<div class="accordion" id="intersection-accordion">')

# Sort intersections by count (descending)
sorted_names <- names(all_intersections_filt)[order(
  sapply(all_intersections_filt, function(x) x$count),
  decreasing = TRUE
)]

  # Generate individual sections for each intersection
for (name in sorted_names) {
  data <- all_intersections_filt[[name]]
        
  prot_map <- ann_df %>%
  filter(pep_seq %in% data$elements) %>%
    group_by(pep_seq) %>%
    summarise(proteins = list(prot_acc),  protein_description = list(prot_desc),.groups = "drop")

    lista_final <- list(
      name = name,
      count = data$count,
      elements = data$elements,
      proteins =   prot_map %>% 
                  mutate(protein_ids = sapply(proteins, paste, collapse = "; ")) %>% 
                  pull(protein_ids),
      protein_description =   prot_map %>%    
                  mutate(protein_description = sapply(protein_description, paste, collapse = "; ")) %>%
                  pull(protein_description) ,
      set_indicators = ntermreport::create_set_indicators(data$sets)
    )
    log_info(name)
    log_info(length(lista_final$proteins))
    log_info(length(lista_final$protein_description))
    log_info(length(lista_final$elements))
    
    ntermreport::render_child(data = lista_final ,
                 label = 'protein' , 
                 path= "",  
                 template='_template_intersection.Rmd') 
  }
    cat('</div>') 


```

#  Overlapping Acetylated Peptide Within Groups

::: panel-tabset
```{r  perGroup}
#| results: asis

#grp_value

for (grp in grp_value){ 

  sample_grp  <- expdesign %>% filter( Group == grp ) %>% pull(Sample)
  cat( '## ', paste(grp,sep=' '), '\n\n')
  log_info(paste('inside -  perGroup ',grp,collapse= '\n'))
  my_list_ <- list()
  for ( g_ in sample_grp) {
        my_list_[[g_]] <- l_ace_[[g_]]$pep_seq
   }

  p <- upsetjs() %>%  fromList(my_list_ )  %>% generateDistinctIntersections()
  print(htmltools::tagList(p))
  cat('\n\n') # Add space before the next tab

}


```

:::